---
title: "Respirometry analysis: Part 1"
subtitle: "Pull in data and format it for use in the `respR` package"
author: "Sarah Tanja"
date: "`r format(Sys.time(), '%d %B, %Y')`"  
format:
  html:
    df-print: paged
    toc: true
    toc-location: right
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-fold: false
    code-tools: true
    code-copy: true
    highlight-style: breeze
    code-overflow: wrap
    theme: minty
editor: 
  markdown: 
    wrap: 72
---

# Install and load packages

```{r}
# Install packages
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr')
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate')
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR')

# Load packages
library(dplyr)
library(tidyverse)
library(lubridate)
library(respR)
```

# About accounting for background

[calc_rate.bg - calculate background
oxygen](https://januarharianto.github.io/respR/articles/closed.html#summary:~:text=calc_rate.bg%20%2D%20calculate%20background%20oxygen)

> The presence of micro-organisms and their oxygen use may be a
> potential source of experimental bias, and we usually want to account
> for background respiration rates during experiments by conducting
> empty or “blank” control experiments with no specimens to quantify it.

# By run

## Respiration

Lets calculate the background for each run in the oxygen consumption
portion of the respirometry test

Make a dataframe of the 'blanks'

```{r}
run_blanks <- resp_log %>% 
  filter(id == "blank") %>% 
  select(id, run, channel)

run_blanks
```

Use it to get the background for each resp_run ~~and remove the blank
from each respiration run dataset~~

```{r}
# run 1
# first calculate the background rate
# This one starts oxygen production at min 25
resp_run_1 <- resp_run_1 %>% filter(delta_t < 25)
bg_insp <- inspect(resp_run_1, time = 1, oxygen = "7")
bg_rate_resp_1 <- calc_rate.bg(bg_insp)
# then remove the background blank from the dataset
#resp_run_1 <- resp_run_1 %>%  select(-"7")

# run 2
bg_insp <- inspect(resp_run_2, time = 1, oxygen = "2")
bg_rate_resp_2 <- calc_rate.bg(bg_insp)
#resp_run_2 <- resp_run_2 %>%  select(-"2")

# 3
# This one starts oxygen production at min 25
resp_run_3 <- resp_run_3 %>% filter (delta_t < 25)
bg_insp <- inspect(resp_run_3, time = 1, oxygen = "4")
bg_rate_resp_3 <- calc_rate.bg(bg_insp)
#resp_run_3 <- resp_run_3 %>%  select(-"4")

# 4
bg_insp <- inspect(resp_run_4, time = 1, oxygen = "1")
bg_rate_resp_4 <- calc_rate.bg(bg_insp)
#resp_run_4 <- resp_run_4 %>%  select(-"1")

# 5
bg_insp <- inspect(resp_run_5, time = 1, oxygen = "3")
bg_rate_resp_5 <- calc_rate.bg(bg_insp)
#resp_run_5 <- resp_run_5 %>%  select(-"3")

# 6
bg_insp <- inspect(resp_run_6, time = 1, oxygen = "5")
bg_rate_resp_6 <- calc_rate.bg(bg_insp)
#resp_run_6 <- resp_run_6 %>%  select(-"5")

# 7
bg_insp <- inspect(resp_run_7, time = 1, oxygen = "7")
bg_rate_resp_7 <- calc_rate.bg(bg_insp)
#resp_run_7 <- resp_run_7 %>%  select(-"7")

# 8
bg_insp <- inspect(resp_run_8, time = 1, oxygen = "5")
bg_rate_resp_8 <- calc_rate.bg(bg_insp)
#resp_run_8 <- resp_run_8 %>%  select(-"5")

# 9
bg_insp <- inspect(resp_run_9, time = 1, oxygen = "4")
bg_rate_resp_9 <- calc_rate.bg(bg_insp)
#resp_run_9 <- resp_run_9 %>%  select(-"4")

# 10
bg_insp <- inspect(resp_run_10, time = 1, oxygen = "3")
bg_rate_resp_10 <- calc_rate.bg(bg_insp)
#resp_run_10 <- resp_run_10 %>%  select(-"3")

# 11
# Bad data at beginning needs to be trimmed
resp_run_11 <- resp_run_11 %>% filter(delta_t > 6)
bg_insp <- inspect(resp_run_11, time = 1, oxygen = "2")
bg_rate_resp_11 <- calc_rate.bg(bg_insp)
#resp_run_11 <- resp_run_11 %>%  select(-"2")
```

### Normality check

```{r}
bg.rates <- c(
  bg_rate_resp_1$summary$rate.bg,
  bg_rate_resp_2$summary$rate.bg,
  bg_rate_resp_3$summary$rate.bg,
  bg_rate_resp_4$summary$rate.bg,
  bg_rate_resp_5$summary$rate.bg,
  bg_rate_resp_6$summary$rate.bg,
  bg_rate_resp_7$summary$rate.bg,
  bg_rate_resp_8$summary$rate.bg,
  bg_rate_resp_9$summary$rate.bg,
  bg_rate_resp_10$summary$rate.bg,
  bg_rate_resp_11$summary$rate.bg)

run <- c(1:11)

resp_run_bg <- tibble(run, bg.rates)

print(resp_run_bg)
```

```{r}
hist(resp_bg$bg.rates)
```

```{r}
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car')
library(car)
qqPlot(resp_bg$bg.rates)
```

## Photosynthesis

The chambers that were blanks are the same

```{r}
# 1
# first calculate the background rate
bg_insp <- inspect(phot_run_1, time = 1, oxygen = "7")
bg_rate_phot_1 <- calc_rate.bg(bg_insp)
# then remove the background blank from the dataset
#phot_run_1 <- phot_run_1 %>%  select(-"7")

# 2
bg_insp <- inspect(phot_run_2, time = 1, oxygen = "2")
bg_rate_phot_2 <- calc_rate.bg(bg_insp)
#phot_run_2 <- phot_run_2 %>%  select(-"2")

# 3
# bad data at beginning
phot_run_3 <- phot_run_3 %>% filter(delta_t > 50)
bg_insp <- inspect(phot_run_3, time = 1, oxygen = "4")
bg_rate_phot_3 <- calc_rate.bg(bg_insp)
#phot_run_3 <- phot_run_3 %>%  select(-"4")

# 4
bg_insp <- inspect(phot_run_4, time = 1, oxygen = "1")
bg_rate_phot_4 <- calc_rate.bg(bg_insp)
#phot_run_4 <- phot_run_4 %>%  select(-"1")

# 5
bg_insp <- inspect(phot_run_5, time = 1, oxygen = "3")
bg_rate_phot_5 <- calc_rate.bg(bg_insp)
#phot_run_5 <- phot_run_5 %>%  select(-"3")

# 6
bg_insp <- inspect(phot_run_6, time = 1, oxygen = "5")
bg_rate_phot_6 <- calc_rate.bg(bg_insp)
#phot_run_6 <- phot_run_6 %>%  select(-"5")

# 7
bg_insp <- inspect(phot_run_7, time = 1, oxygen = "7")
bg_rate_phot_7 <- calc_rate.bg(bg_insp)
#phot_run_7 <- phot_run_7 %>%  select(-"7")

# 8
# bad data at end
phot_run_8 <- phot_run_8 %>% filter(delta_t < 53)
bg_insp <- inspect(phot_run_8, time = 1, oxygen = "5")
bg_rate_phot_8 <- calc_rate.bg(bg_insp)
#phot_run_8 <- phot_run_8 %>%  select(-"5")

# 9
bg_insp <- inspect(phot_run_9, time = 1, oxygen = "4")
bg_rate_phot_9 <- calc_rate.bg(bg_insp)
#phot_run_9 <- phot_run_9 %>%  select(-"4")

# 10
bg_insp <- inspect(phot_run_10, time = 1, oxygen = "3")
bg_rate_phot_10 <- calc_rate.bg(bg_insp)
#phot_run_10 <- phot_run_10 %>%  select(-"3")

# 11
bg_insp <- inspect(phot_run_11, time = 1, oxygen = "2")
bg_rate_phot_11 <- calc_rate.bg(bg_insp)
#phot_run_11 <- phot_run_11 %>%  select(-"2")

```

### Normality check

```{r}
phot.bg.rates <- c(
  bg_rate_phot_1$summary$rate.bg,
  bg_rate_phot_2$summary$rate.bg,
  bg_rate_phot_3$summary$rate.bg,
  bg_rate_phot_4$summary$rate.bg,
  bg_rate_phot_5$summary$rate.bg,
  bg_rate_phot_6$summary$rate.bg,
  bg_rate_phot_7$summary$rate.bg,
  bg_rate_phot_8$summary$rate.bg,
  bg_rate_phot_9$summary$rate.bg,
  bg_rate_phot_10$summary$rate.bg,
  bg_rate_phot_11$summary$rate.bg)

run <- c(1:11)

phot_run_bg <- tibble(run, phot.bg.rates)

print(phot_run_bg)
```

```{r}
hist(phot_bg$phot.bg.rates)
```

```{r}
qqPlot(phot_bg$phot.bg.rates)
```

# By channel

```{r}
channel_blanks <- resp_log %>% 
  filter(id == "blank") %>% 
  group_by(channel) %>% 
  summarize(run = list(run))

# Print the result
for (i in 1:nrow(channel_blanks)) {
  cat("Channels used to blank runs (channel : run)", channel_blanks$channel[i], ":", unlist(channel_blanks$run[[i]]), "\n")
}

```

## Respiration

```{r}
# channel 1
# channel 1 was the blank for run 4
bg_insp <- inspect(resp_run_4, time = 1, oxygen = "1")
bg_rate_resp_channel_1 <- calc_rate.bg(bg_insp)

# channel 2
resp_channel_2 <- resp_run_2 %>% 
  select (delta_t, "2") %>% 
  full_join(resp_run_11 %>% select(delta_t, "2"), by = "delta_t") %>% 
  arrange(delta_t)

bg_insp <- inspect(resp_channel_2, time = 1, oxygen = 2:3)
bg_rate_resp_channel_2 <- calc_rate.bg(bg_insp)

# channel 3
resp_channel_3 <- resp_run_5 %>% 
  select (delta_t, "3") %>% 
  full_join(resp_run_10 %>% select(delta_t, "3"), by = "delta_t") %>% 
  arrange(delta_t)

bg_insp <- inspect(resp_channel_3, time = 1, oxygen = 2:3)
bg_rate_resp_channel_3 <- calc_rate.bg(bg_insp)

# channel 4
resp_channel_4 <- resp_run_3 %>% 
  select (delta_t, "4") %>% 
  full_join(resp_run_9 %>% select(delta_t, "4"), by = "delta_t") %>% 
  arrange(delta_t)

bg_insp <- inspect(resp_channel_4, time = 1, oxygen = 2:3)
bg_rate_resp_channel_4 <- calc_rate.bg(bg_insp)

# channel 5
# channel 5 was the blank for runs 6 & 8
resp_channel_5 <- resp_run_6 %>% 
  select (delta_t, "5") %>% 
  full_join(resp_run_8 %>% select(delta_t, "5"), by = "delta_t") %>% 
  arrange(delta_t)

bg_insp <- inspect(resp_channel_5, time = 1, oxygen = 2:3)
bg_rate_resp_channel_5 <- calc_rate.bg(bg_insp)

# channel 7
# channel 7 was the blank for runs 1 & 7
resp_channel_7 <- resp_run_1 %>% 
  select (delta_t, "7") %>% 
  full_join(resp_run_7 %>% select(delta_t, "7"), by = "delta_t") %>% 
  arrange(delta_t)

bg_insp <- inspect(resp_channel_7, time = 1, oxygen = 2:3)
bg_rate_resp_channel_7 <- calc_rate.bg(bg_insp)

```

```{r}

```

### Normality check

```{r}
bg.rates.channel <- c(
  bg_rate_resp_channel_1[["rate.bg.mean"]],
  bg_rate_resp_channel_2[["rate.bg.mean"]],
  bg_rate_resp_channel_3[["rate.bg.mean"]],
  bg_rate_resp_channel_4[["rate.bg.mean"]],
  bg_rate_resp_channel_5[["rate.bg.mean"]],
  bg_rate_resp_channel_7[["rate.bg.mean"]]
  )

channel <- c(1:5, 7)

bg_resp_channel <- tibble(channel, bg.rates.channel)

print(bg_resp_channel)
```

```{r}
hist(bg_resp_channel$bg.rates.channel)
```

## Photosynthesis

# By mean
