---
title: "Respirometry analysis: Part 2"
subtitle: "Get background rates from the `blank` chambers"
author: "Sarah Tanja"
date: "`r format(Sys.time(), '%d %B, %Y')`"  
format:
  html:
    df-print: paged
    toc: true
    toc-location: right
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-fold: false
    code-tools: true
    code-copy: true
    highlight-style: breeze
    code-overflow: wrap
    theme: minty
editor: 
  markdown: 
    wrap: 72
---

# Install and load packages

```{r messages = FALSE}
# Install packages
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr')
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate')
if ("respR" %in% rownames(installed.packages()) == 'FALSE') install.packages('respR')
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car')
if ("kableExtra" %in% rownames(installed.packages()) == 'FALSE') install.packages('kableExtra')

# Load packages
library(dplyr)
library(tidyverse)
library(lubridate)
library(respR)
library(car)
library(kableExtra)
```

# Load in dataframes

```{r}
# Get a list of all CSV files in your output folder
csv_files <- list.files("output", pattern = "\\.csv$", full.names = TRUE)

# Loop through each CSV file and read it into a dataframe
for (file in csv_files) {
  # Extract the dataframe name from the file name
  df_name <- sub("\\.csv$", "", basename(file))
  # Read the CSV file into a dataframe and assign it the dataframe name
  assign(df_name, read.csv(file, check.names = FALSE))
}

# Now you'll have dataframes named after the CSV files
# For example, if you have a file named "example.csv", you'll have a dataframe named "example"
```

# Respiration QAQC

## Join resp_rates to merged_metadata

```{r}
resp_rates <- left_join(resp_rates, merged_metadata)
```

## bad data

```{r}
# filter for rates that are positive
resp_badrates <- resp_rates %>% filter(resp_rate > 0 )

print(resp_badrates)
```

```{r}
# filter for r2 values that are below 0.8
resp_badr2 <- resp_rates %>% filter(resp_r2 < 0.8)
print(resp_badr2)
```

```{r}
resp_baddies <- full_join(resp_badr2, resp_badrates)

# Print a pretty table
resp_baddies %>%
  select(run, channel) %>%
  arrange(run, channel) %>% 
  kable(caption = "Respiration Trouble", format = "html", 
        col.names = c("Run", "Channel")) %>%
  kable_styling(full_width = FALSE)
```

Inspect the 'bad data' individually

#### 1.7

```{r}
resp_run_1 %>% select('delta_t', '7') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '7') %>%  auto_rate()
```

```{r}
new1.7 <-  resp_run_1 %>%
  inspect(time = 'delta_t', oxygen = '7') %>%
  calc_rate(from = 4, to = 15, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new1.7
this.run <- 1
this.channel <- 7
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, NA),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, NA)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2")
```

#### 2.2

```{r}
resp_run_2 %>% select('delta_t', '2') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '2') %>%  auto_rate()
```

```{r}
new2.2 <-  resp_run_2 %>%
  inspect(time = 'delta_t', oxygen = '2') %>%
  calc_rate(from = 0, to = 6, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new2.2
this.run <- 2
this.channel <- 2
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2")
```

#### 2.3

```{r}
resp_run_2 %>% select('delta_t', '3') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '3') %>%  auto_rate()
```

```{r}
new2.3 <-  resp_run_2 %>%
  inspect(time = 'delta_t', oxygen = '3') %>%
  calc_rate(from = 0, to = 5, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new2.3
this.run <- 2
this.channel <- 3
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 2.4

```{r}
resp_run_2 %>% select('delta_t', '4') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '4') %>%  auto_rate()
```

```{r}
new2.4 <-  resp_run_2 %>%
  inspect(time = 'delta_t', oxygen = '4') %>%
  calc_rate(from = 0, to = 7, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new2.4
this.run <- 2
this.channel <- 4
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 6.3

```{r}
resp_run_6 %>% select('delta_t', '3') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '3') %>%  auto_rate()
```

```{r}
new6.3 <-  resp_run_6 %>%
  inspect(time = 'delta_t', oxygen = '3') %>%
  calc_rate(from = 0, to = 15, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new6.3
this.run <- 6
this.channel <- 3
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 6.7

```{r}
resp_run_6 %>% select('delta_t', '7') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '7') %>%  auto_rate()
```

```{r}
new6.7 <-  resp_run_6 %>%
  inspect(time = 'delta_t', oxygen = '7') %>%
  calc_rate(from = 0, to = 12, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new6.7
this.run <- 6
this.channel <- 7
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 10.1

```{r}
resp_run_10 %>% select('delta_t', '1') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '1') %>%  auto_rate()
```

```{r}
new10.1 <-  resp_run_10 %>%
  inspect(time = 'delta_t', oxygen = '1') %>%
  calc_rate(from = 0, to = 12, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new10.1
this.run <- 10
this.channel <- 1
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 10.2

```{r}
resp_run_10 %>% select('delta_t', '2') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '2') %>%  auto_rate()
```

```{r}
new10.2 <-  resp_run_10 %>%
  inspect(time = 'delta_t', oxygen = '2') %>%
  calc_rate(from = 0, to = 12, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new10.2
this.run <- 10
this.channel <- 2
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 10.4

```{r}
resp_run_10 %>% select('delta_t', '4') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '4') %>%  auto_rate()
```

```{r}
new10.4 <-  resp_run_10 %>%
  inspect(time = 'delta_t', oxygen = '4') %>%
  calc_rate(from = 0, to = 12, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new10.4
this.run <- 10
this.channel <- 4
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 10.5

```{r}
resp_run_10 %>% select('delta_t', '5') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '5') %>%  auto_rate()
```

```{r}
new10.5 <-  resp_run_10 %>%
  inspect(time = 'delta_t', oxygen = '5') %>%
  calc_rate(from = 0, to = 10, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new10.5
this.run <- 10
this.channel <- 5
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 10.7

```{r}
resp_run_10 %>% select('delta_t', '7') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '7') %>%  auto_rate()
```

```{r}
new10.7 <-  resp_run_10 %>%
  inspect(time = 'delta_t', oxygen = '7') %>%
  calc_rate(from = 0, to = 10, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new10.7
this.run <- 10
this.channel <- 7
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```

#### 11.4

```{r}
resp_run_11 %>% select('delta_t', '4') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '4') %>%  auto_rate()
```

```{r}
new11.4 <-  resp_run_11 %>%
  inspect(time = 'delta_t', oxygen = '4') %>%
  calc_rate(from = 25, to = 30, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new11.4
this.run <- 11
this.channel <- 4
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_resp_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
resp_rates <- resp_rates %>%
  mutate(
    qaqc_resp_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_resp_rate),
    qaqc_resp_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_resp_r2),
    qaqc_resp_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_resp_bg_adj_rate)
  )



filter(resp_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_resp_rate", "qaqc_resp_r2", "qaqc_resp_bg_adj_rate")
```
## Save updates
```{r}
write.csv(resp_rates)
```


# Photosynthesis QAQC

## Join phot_rates to merged_metadata

```{r}
phot_rates <- left_join(phot_rates, merged_metadata)
```

## bad data

```{r}
# Look for rates that are negative
phot_badrates <- phot_rates %>% filter(phot_rate<0)
print(phot_badrates)
```

```{r}
# Look for rate r2 values that are less than 0.8
phot_badr2 <- phot_rates %>% filter(phot_r2<0.8)
print(phot_badr2)
```

```{r}
phot_baddies <- full_join(phot_badr2, phot_badrates)

# Print a pretty table
phot_baddies %>%
  select(run, channel) %>%
  arrange(run, channel) %>% 
  kable(caption = "Photosynthesis Trouble", format = "html", 
        col.names = c("Run", "Channel")) %>%
  kable_styling(full_width = FALSE)
```

### 1.7

```{r}
phot_run_1 %>% select('delta_t', '7') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '7') %>%  auto_rate()
```

```{r}
new1.7 <-  phot_run_1 %>%
  inspect(time = 'delta_t', oxygen = '7') %>%
  calc_rate(from = 41, to = 45, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new1.7
this.run <- 1
this.channel <- 7
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, NA),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, NA),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, NA)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 2.1

```{r}
phot_run_2 %>% select('delta_t', '1') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '1') %>%  auto_rate()
```

```{r}
new2.1 <-  phot_run_2 %>%
  inspect(time = 'delta_t', oxygen = '1') %>%
  calc_rate(from = 43, to = 50, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new2.1
this.run <- 2
this.channel <- 1
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 2.3

```{r}
phot_run_2 %>% select('delta_t', '3') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '3') %>%  auto_rate()
```

```{r}
new2.3 <-  phot_run_2 %>%
  inspect(time = 'delta_t', oxygen = '3') %>%
  calc_rate(from = 43, to = 50, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new2.3
this.run <- 2
this.channel <- 3
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 3.3

```{r}
phot_run_3 %>% select('delta_t', '3') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '3') %>%  auto_rate()
```

```{r}
new3.3 <-  phot_run_3 %>%
  inspect(time = 'delta_t', oxygen = '3') %>%
  calc_rate(from = 54, to = 56, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new3.3
this.run <- 3
this.channel <- 3
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 4.2

```{r}
phot_run_4 %>% select('delta_t', '2') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '2') %>%  auto_rate()
```

```{r}
new4.2 <-  phot_run_4 %>%
  inspect(time = 'delta_t', oxygen = '2') %>%
  calc_rate(from = 38, to = 44, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new4.2
this.run <- 4
this.channel <- 2
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 5.5

```{r}
phot_run_5 %>% select('delta_t', '5') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '5') %>%  auto_rate()
```

```{r}
new5.5 <-  phot_run_5 %>%
  inspect(time = 'delta_t', oxygen = '5') %>%
  calc_rate(from = 40, to = 55, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new5.5
this.run <- 5
this.channel <- 5
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 7.1

```{r}
phot_run_7 %>% select('delta_t', '1') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '1') %>%  auto_rate()
```

```{r}
new7.1 <-  phot_run_7 %>%
  inspect(time = 'delta_t', oxygen = '1') %>%
  calc_rate(from = 40, to = 50, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new7.1
this.run <- 7
this.channel <- 1
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 7.3

```{r}
phot_run_7 %>% select('delta_t', '3') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '3') %>%  auto_rate()
```

```{r}
new7.3 <-  phot_run_7 %>%
  inspect(time = 'delta_t', oxygen = '1') %>%
  calc_rate(from = 52, to = 60, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new7.3
this.run <- 7
this.channel <- 3
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 7.4

```{r}
phot_run_7 %>% select('delta_t', '4') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '4') %>%  auto_rate()
```

```{r}
new7.4 <-  phot_run_7 %>%
  inspect(time = 'delta_t', oxygen = '4') %>%
  calc_rate(from = 52, to = 60, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new7.4
this.run <- 7
this.channel <- 4
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 8.2

```{r}
phot_run_8 %>% select('delta_t', '2') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '2') %>%  auto_rate()
```

::: callout
This data is showing an entirely negative slope in the light phase, this
may indicate that either the data is bad altogether, or that the
symbionts of this anemone were not doing photosynthesis.
:::

```{r}
new8.2 <-  phot_run_8 %>%
  inspect(time = 'delta_t', oxygen = '2') %>%
  calc_rate(from = 42, to = 52, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new8.2
this.run <- 8
this.channel <- 2
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 11.1

```{r}
phot_run_11 %>% select('delta_t', '1') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '1') %>%  auto_rate()
```

```{r}
new11.1 <-  phot_run_11 %>%
  inspect(time = 'delta_t', oxygen = '1') %>%
  calc_rate(from = 54, to = 58, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new11.1
this.run <- 11
this.channel <- 1
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```

### 11.3

```{r}
phot_run_11 %>% select('delta_t', '3') %>% drop_na() %>% 
inspect(time = 'delta_t', oxygen = '3') %>%  auto_rate()
```

```{r}
new11.3 <-  phot_run_11 %>%
  inspect(time = 'delta_t', oxygen = '3') %>%
  calc_rate(from = 52, to = 59, by = 'time')
```

```{r}
### ONLY CHANGE THESE VALUES ###
new.calc <- new11.3
this.run <- 11
this.channel <- 3
################################

new.rate <- new.calc$summary$rate[1]
new.rsq <- new.calc$summary$rsq[1]

blank <- paste0("bg_rate_phot_", this.run)
# new background adjusted object
new.adj <- adjust_rate(new.calc, by = get(blank)) 
# from the background-adjusted object, pull the adjusted rate
new.bg.adj.rate <- new.adj$summary$rate.adjusted[1]

# Add qaqc value to new column based on conditions in 'run' and 'channel' columns
phot_rates <- phot_rates %>%
  mutate(
    qaqc_phot_rate = ifelse(run == this.run &
                         channel == this.channel, new.rate, qaqc_phot_rate),
    qaqc_phot_r2 = ifelse(run == this.run &
                       channel == this.channel, new.rsq, qaqc_phot_r2),
    qaqc_phot_bg_adj_rate = ifelse(run == this.run &
                       channel == this.channel, new.bg.adj.rate, qaqc_phot_bg_adj_rate)
  )



filter(phot_rates, run == this.run & channel == this.channel) %>% select("run", "channel", "treatment", "qaqc_phot_rate", "qaqc_phot_r2", "qaqc_phot_bg_adj_rate")
```
