# Document for Anova analysis check the git link for the tutorial

# ANOVA notes with input

what is it? stat test to see if two or more pop means are different.

primarily used to compare 3 or more groups

# Test we are using is one-way anova specifically

If the between variance is significantly larger than the within variance, the group means are declared to be different

variance comparison done using the ratio (variance between/ variance within)

after ratio is complete compare to fisher probability threshold (F-test)

How to do (example using the penguins data set):

#install.packages("palmerpenguins")

#library(palmerpenguins)

#library(tidyverse) \# dat \<- penguins %\>% \# select(species, flipper_length_mm)

#replace this with the resp or Photosynth data \#

note: the pipe operator (%\>%) takes the output of the expression on its left and passes it as the first argument to the function on its right

#summary(dat)

Aim of ANOVA: #study whether measurements are similar across different modalities (also called levels or treatments in the context of ANOVA) of a categorical variable #compare the impact of the different levels of a categorical variable on a quantitative variable #explain a quantitative variable based on a qualitative variable

# Null and Alternative hypothesis

H0: 1=2=3 (equal across the data sets)

H1: AT LEAST one is different from the other two (could be any of the lot)

large sample is n \>= 30

if normality is assumed, test the homogeneity of variances #variances equal, ANOVA #variances not equal, Welch ANOVA #normality not assumed, Kruskal-Wallis test

## Direct implication:

Anemone treatments are the variables

# How to do it
#install tidyverse, dplyr, car and load 

```{r}
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr')
if ("car" %in% rownames(installed.packages()) == 'FALSE') install.packages('car')

library(dplyr)
library(tidyverse)
library(car)
```


```{r}
ANOVA_RATES <- read.csv('rates.csv')

```

```{r}

```

