---
title: "Respirometry analysis: Part 1"
subtitle: "Pull in data and format it for use in the `respR` package"
author: "Sarah Tanja"
date: "`r format(Sys.time(), '%d %B, %Y')`"  
format:
  html:
    df-print: paged
    toc: true
    toc-location: right
    smooth-scroll: true
    link-external-icon: true
    link-external-newwindow: true
    code-fold: false
    code-tools: true
    code-copy: true
    highlight-style: breeze
    code-overflow: wrap
    theme: minty
editor: 
  markdown: 
    wrap: 72
---

# Install and load packages

```{r}
# Install packages
if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages('tidyverse')
if ("dplyr" %in% rownames(installed.packages()) == 'FALSE') install.packages('dplyr')
if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages('lubridate')
if ("stringr" %in% rownames(installed.packages()) == 'FALSE') install.packages('stringr')

# Load packages
library(dplyr)
library(tidyverse)
library(lubridate)
library(stringr)
```

# Pull in metadata

```{r}
nem_metadata <- read_csv("../metadata/nem_metadata.csv")
```

# Pull in data

```{r}
rlc <- read_csv("data/ALL55_anemones_RLSs_26APR2024.csv")
```

# Filter data

```{r}
# Join metadata with data by matching "No." to "RLC_start"
rlc_mer <- left_join(rlc, nem_metadata, by = c("No." = "RLC_start")) 

# Bring the "RLC_start" column back in by re-joining to the metadata
rlc_merged <- left_join(rlc_mer, nem_metadata) 

rlc_merged <- rlc_merged %>% relocate (RLC_start, .before = RLC_end)
```

# Fill data

```{r}

rlc_fill <- rlc_merged %>% 
  filter(Type %in% c("REG2", "FO")) %>% 
  fill(treatment, .direction = "up") %>% 
  fill(id, .direction = "up") %>% 
  fill(No., .direction = "up")%>% 
  fill("1:Fv/Fm", .direction = "up") %>% 
  filter(No. == lead(RLC_start)) %>% 
  select(No., "1:F","1:Fv/Fm", id, treatment )

rlc_fill
```

## Check that there are 55 rows, with 11 treatments that each have 5 replicates of each

```{r}
# Get unique treatments and count occurrences of each
table(rlc_fill$treatment) # there should be 5 of each
table(rlc_fill$id) # there should be 1 of each
```

## Extract alpha, ETRm, & Ik values

```{r}
# Define a regex pattern to extract alpha, ETRm, and Ik values
regex <- "alpha: ([\\d.-]+), ETRm: ([\\d.-]+), Ik: ([\\d.-]+)"

# Extract the rows that match the pattern
photobot_results <- rlc_fill %>%
  mutate(match = str_extract(`1:F`, regex)) %>%
  filter(!is.na(match)) %>%
  mutate(alpha = str_extract(match, "(?<=alpha: )[\\d.-]+"),
         ETRm = str_extract(match, "(?<=ETRm: )[\\d.-]+"),
         Ik = str_extract(match, "(?<=Ik: )[\\d.-]+")) %>%
  select(id, treatment, "1:Fv/Fm", alpha, ETRm, Ik)

# rename FvFm & make numeric
photobot_results <- photobot_results %>% 
  rename(FvFm = "1:Fv/Fm") 

photobot_results <- photobot_results %>% 
  mutate(alpha = as.numeric(alpha)) %>% 
  mutate(ETRm = as.numeric(ETRm)) %>% 
  mutate(Ik = as.numeric(Ik))

print(photobot_results)

```

# Make treatment a factor

```{r}
photobot_results <- photobot_results %>% 
  mutate(treatment = as.factor(treatment))

class(photobot_results$treatment)
```

# Fv/Fm

```{r}
fvfm_aov <- aov(FvFm ~ treatment,
  data = photobot_results
)

summary(fvfm_aov)
```

```{r}

FvFm <- ggplot(photobot_results) +
  aes(x = treatment, y = FvFm, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Photosynthetic efficiency (Fv/Fm)",
    x = "Leachate concentration (mg/L)",
    y = "Photosynthetic efficiency (Fv/Fm)"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4")) 

FvFm

# Save the ggplot object to a PNG file
ggsave(filename = "FvFm.png", plot = FvFm, width = 11, height = 6, dpi = 600)

```

# Alpha

## plot

```{r}
Alpha <- ggplot(photobot_results) +
  aes(x = treatment, y = alpha, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Alpha",
    x = "Leachate concentration (mg/L)",
    y = "Alpha"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4")) 

Alpha

# Save the ggplot object to a PNG file
ggsave(filename = "Alpha.png", plot = Alpha, width = 11, height = 6, dpi = 300)
```

## anova

```{r}
alpha_aov <- aov(alpha ~ treatment,
  data = photobot_results
)

summary(alpha_aov)
```

## PVC only

Filter the dataframe for just the PVC and control treatments

```{r}
photobot_pvc <- photobot_results %>% 
  filter(treatment %in% c('Control', 'PVC_0.01_mgL', 'PVC_0.1_mgL', 'PVC_1_mgL', 'PVC_10_mgL', 'PVC_100_mgL'))

photobot_pvc
```

### plot

```{r}
PVCAlpha <- ggplot(photobot_pvc) +
  aes(x = treatment, y = alpha, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Alpha",
    x = "Leachate concentration (mg/L)",
    y = "Alpha"
  ) +
  scale_x_discrete(labels = c("Control ", "PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4")) 

PVCAlpha

# Save the ggplot object to a PNG file
ggsave(filename = "PVCAlpha.png", plot = PVCAlpha, width = 11, height = 6, dpi = 300)

```

Alpha (α) Definition: Alpha (α) is defined as the initial slope of the
Rapid Light Curve at low light intensities. It represents the
photosynthetic efficiency, indicating how effectively a photosynthetic
organism can use available light for photosynthesis under low light
conditions. A higher alpha value indicates that the photosynthetic
organism is highly efficient at utilizing low light levels, which can be
crucial for survival in shaded or low-light environments. Conversely, a
lower alpha suggests less efficiency in low light conditions. The alpha
value is calculated from the initial linear portion of the Rapid Light
Curve. It is determined by fitting a straight line to the data points at
the lowest light intensities and calculating the slope of this line. The
steeper the slope, the higher the photosynthetic efficiency at low
light.

### anova

```{r}
alpha_pvc_aov <- aov(alpha ~ treatment,
  data = photobot_pvc
)

summary(alpha_pvc_aov)
```

### Dunnett's test

Is the lowest PVC leachate concentration actually larger than the
control? Or is this a `Type II error`, or error of multiple tests? I
followed the [stats and r blog to conduct a post-hoc Dunnett's
test](https://statsandr.com/blog/anova-in-r/#post-hoc-test:~:text=plot(TukeyHSD(res_aov))-,Dunnett%E2%80%99s%20test,-We%20have%20seen)
on the above ANOVA result.

```{r}
if ("multcomp" %in% rownames(installed.packages()) == 'FALSE') install.packages('multcomp')
library(multcomp)

# Make sure `treatment` is a factor


# Dunnett's test:
post_test <- glht(alpha_pvc_aov,
  linfct = mcp(treatment = "Dunnett")
)

summary(post_test)
```

::: callout-warning
The Dunnett's test adjusted *p* value comparing the lowest PVC
concentration to the control (*p* = .0671) indicated that the treatment
is not significantly different from the control.
:::

::: callout-note
Note that in R, by default, the reference category for a [factor
variable](https://statsandr.com/blog/data-types-in-r/#factor) is the
first category in alphabetical order. This is the reason that, by
default, the reference species is `Control`.

The reference category can be changed with the `relevel()` function (or
with the [`{questionr}`
addin](https://statsandr.com/blog/rstudio-addins-or-how-to-make-your-coding-life-easier/#reordering-factors)).
Look
[here](https://statsandr.com/blog/anova-in-r/#post-hoc-test:~:text=%23%20Change%20reference%20category%3A)
for an example.
:::

Examining the 95% family-wise confidence level comparing the treatments
to the controls:

```{r}

# Open the PNG device
png(filename = "Dunnett.png", width = 800, height = 600, res = 72)

### use sufficiently large upper margin
old.par <- par(mai=c(1,2.25,1.25,0.25), no.readonly=TRUE)
### plot
plot(post_test)
par(old.par)

# Close the device
dev.off()

```

```{r}
### use sufficiently large upper margin
old.par <- par(mai=c(1,2.25,1.25,0.25))
### plot
plot(post_test)
par(old.par)
```

## PP only?

```{r}
photobot_pp <- photobot_results %>% 
  filter(treatment %in% c('Control', 'PP_A_0.01_mgL', 'PP_A_0.1_mgL', 'PP_A_1_mgL', 'PP_A_10_mgL', 'PP_A_100_mgL'))

photobot_pp
```

```{r}
alpha_pp_aov <- aov(alpha ~ treatment,
  data = photobot_pp
)

summary(alpha_pp_aov)
```

```{r}
ggplot(photobot_pp) +
  aes(x = treatment, y = alpha, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. Alpha",
    x = "Leachate concentration (mg/L)",
    y = "Alpha"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4"))
```

# ETRm

```{r}
etrm_aov <- aov(ETRm ~ treatment,
  data = photobot_results
)

summary(etrm_aov)
```

```{r}
ggplot(photobot_results) +
  aes(x = treatment, y = ETRm, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Concentration vs. rETRm",
    x = "Leachate concentration (mg/L)",
    y = "rETRm"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4"))
```

# Ik

```{r}
ik_aov <- aov(Ik ~ treatment,
  data = photobot_results
)

summary(ik_aov)
```

```{r}
Ik <- ggplot(photobot_results) +
  aes(x = treatment, y = Ik, fill= treatment) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   labs(
    title = "Leachate concentration vs. Saturating Irradiance (Ik)",
    x = "Leachate concentration (mg/L)",
    y = "Saturating Irradiance (PAR umol/m2/s)"
  ) +
  scale_x_discrete(labels = c("Control ", "PPa 0.01", "PPa 0.1","PPa 1", "PPa 10", "PPa 100","PVC 0.01", "PVC 0.1", "PVC 1", "PVC 10", "PVC 100")) +  
  scale_fill_manual(values = c("seashell", "coral", "coral1","coral2","coral3", "coral4","dodgerblue","dodgerblue1","dodgerblue2", "dodgerblue3","dodgerblue4"))

Ik

```

What the hell is wrong with the control!?
