---
title: "Artemia"
author: "Sarah Tanja"
date: '03/27/2024'
categories: [protocols, anemones]
draft: true
toc: true
toc-title: Contents <i class="bi bi-bookmark-heart"></i>
toc-depth: 5
toc-location: left
reference-location: margin
citation-location: margin
link-external-icon: true
link-external-newwindow: true
---


```{r echo=TRUE, message=FALSE, warning=FALSE}
## install packages
if (!require("tidyverse", character.only = TRUE)) { install.packages("tidyverse")}
if (!require("kableExtra", character.only = TRUE)) { install.packages("kableExtra")}
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## load libraries
library(tidyverse)
library(kableExtra)
```

# 27MAR2024

Nauplii counting

Well-plate

```{r}
# Well plate counts of freshly hatched artemia nauplii
A <- c(4,4,6,7,21,11)
B <- c(3,6,8,7,5,9)
C <- c(15,3,13,5,7,14)
D <- c(8,13,8,16,12,5)

# Sum each vector, and take the average
ml_nauplii <- mean(c(sum(A), sum(B), sum(C), sum(D)))
print(ml_nauplii)
```

```{r}
# Multiply it out by volume
# Each anemone gets 4mL of nauplii
ml_nauplii*4
```

## 29-MAR-2024

```{r}
# Well plate counts of freshly hatched artemia nauplii
# counts represent 10uL volumes
A <- c(6,6,10,7,3,11)
B <- c(10,6,14,14,14,9)
C <- c(7,11,12,21,10,13)
D <- c(8,14,11,8,6,13)

naup <- data.frame(c(A, B, C, D)) %>%
  rename(counts = 1)

summary(naup)
sd(naup$counts)

# Calculate coefficient of variation
cv <- sd(naup$counts) / mean(naup$counts) * 100

# Print result
print(cv)

```

So in 10uL we have 10 +/- 4 artemia... We have to multiply 10uL by 400 to get 4mL(4000uL) In 4mL we could have anywhere from $6*400 = 2,400$ to $10*400 = 4,000$ artemia So, not ideal... because that means some anemones would be possibly getting twice as much as others. Sigh. Fuck. Stats.

So I need a better way to consistently count the artemia... so that I'm getting a more accurate count/method of distributing equal numbers. SHIT.

Ok going to use a stir plate and bar and try again

```{r}
# Well plate counts of freshly hatched artemia nauplii
# counts represent 10uL volumes
A <- c(7,7,12,10,14,7)
B <- c(3,16,11,8,8,8)
C <- c(5,5,13,10,11,17)
D <- c(7,14,4,9,8,10)

naup <- data.frame(c(A, B, C, D)) %>%
  rename(counts = 1)

summary(naup)
sd(naup$counts)

# Calculate coefficient of variation
cv <- sd(naup$counts) / mean(naup$counts) * 100

# Print result
print(cv)

```

Same result with stir bar as compared to hand stirring. Well shit, what next!? But I don't want to count a larger volume that will take more time Fuck okay we'll go up to 20uL

```{r}
# Well plate counts of freshly hatched artemia nauplii
# counts represent 20uL volumes
A <- c(18,24,25,30,31,25)
B <- c(32,27,29,27,24,19)
C <- c(24,31,34,21,27,26)
D <- c(31,28,30,33,26,28)

naup <- data.frame(c(A, B, C, D)) %>%
  rename(counts = 1)

summary(naup)
sd(naup$counts)

# Calculate coefficient of variation
cv <- sd(naup$counts) / mean(naup$counts) * 100

# Print result
print(cv)

```

Whoo! Cut coefficient of variation in half baby! I'll fuckking take it.

So. 20uL to 4mL, multiply by 200! 27+/-4 nauplii in 20uL, then there will be anywhere from $24*200=4,800$ and $31*200=6,200$ nauplii fed to each anemone. Good enough. If I wanted it to be narrower I'd probably have to increase the volume and therefore the time and difficulty of counting. I could do a whole analysis on the ideal volume of shrimp counting to minimize coefficient of variation and make really pretty plots! But no... I shan't be tempted...

I CAN COUNT SHRIMP.

## 11-MAR-2024
```{r}
# Well plate counts of freshly hatched artemia nauplii
# counts represent 20uL volumes
A <- c(32,23,20,22,34,21)
B <- c(33,40,21,22,30,31)
C <- c(36,25,34,35,26,33)
D <- c(36,34,22,24,32,32)

naup <- data.frame(c(A, B, C, D)) %>%
  rename(counts = 1)

summary(naup)
sd(naup$counts)

# Calculate coefficient of variation
cv <- sd(naup$counts) / mean(naup$counts) * 100

# Print result
print(cv)

# Max number of artemia in 1mL
(mean(naup$counts)+sd(naup$counts))*50

# Min artemia in 1mL
(mean(naup$counts)-sd(naup$counts))*50

```
So between ~1.7K and ~1.1k artemia each

I think 4mL last time was too much...  This time we do 1mL each instead? When I gave them 4mL at this concentration them last time there was a lot of leftover dead nauplii that just fouled up the dishes...
For this feeding I gave them 1mL each at 15:30 on Thursday April 11th, after lowering the water level to just below the jar lids. I left the water level low until ___
